Supplemental Materials for ``Universals of word order result from optimization of grammars for efficient communication''

### 2019-06-27 writeup/pnas-word-order-si.tex
SI
si

\section{Formalizing Comunicative Efficiency}
\subsection{Relation to Other Work}

@@

Here we discuss how our formalization of communicative efficiency relates to formalizations that have been proposed in the information-theoretic literature on language.
Across the literature, the core idea is to maximize the \key{amount of information} that linguistic forms provide about meanings, while constraining \key{complexity and diversity} of forms, with some differences in the precise formalization of these two quantities \cite{ferreri2003least,ferrericancho2007global,kemp2012kinship,frank2012predicting,goodman2013knowledge,kao2014nonliteral,xu2014numeral,kirby2015compression,regier2015word,xu2016historical,futrell2017memory,zaslavsky2018efficient,bennett2018extremely, peloquin2019interactions,zaslavsky2019semantic}.

The basis for our precise formalization is the function proposed in \cite{ferreri2003least,ferrericancho2007global,futrell2017memory,peloquin2019interactions} as a general efficiency metric for communicative systems.
%The key idea is to maximize the \key{amount of information} that linguistic forms provide about meanings, while constraining \key{complexity and diversity} of forms. %; there are some differences in the formalization of complexity.
If $S$ denotes signals (e.g., words, sentences) and $R$ denotes their referents (e.g., objects in a reference game), then this efficiency metric takes the form (notation slightly varies across these publications):
\begin{equation}\label{eq:eff-general}
I[S, R] - \lambda \cdot H[S]
\end{equation}
where $I[S, R]$ describes the \key{informativity} of the signals $S$ about their referents $R$, and $H[S]$ describes the \key{complexity} of the communication system, and $\lambda > 0$ trades off the two aspects of efficiency.
While prior studies \cite{ferreri2003least, kemp2012kinship,regier2015word,zaslavsky2018efficient} mostly considered settings where the signals $S$ are individual words without further structure, the signals are entire sentences $\utterance$ in our setting.
The underlying messages $R$ which the speaker aims to convey are the syntactic structures $\tree$.
By the principle of compositionality \cite{frege1892sinn}, the meaning of a sentence is a function of the meanings of the parts and how they are combined.
The syntactic structure specifies how the meanings of words are combined; therefore, recovering the syntactic structure is a prerequisite to understanding a sentence correctly.
%The idea that efficient languages will permit the listener to infer grammati
Hence, substituting utterances $\utterance$ for signals $S$, and syntactic structures $\tree$ for underlying messages $R$, into (\ref{eq:eff-general}), we  arrive at the following efficiency metric for word order:
\begin{equation}\label{eq:efficiency-derived}
	R_{\textit{Eff}} := R_{\textit{Pars}} + \lambda \cdot R_\textit{Pred}
\end{equation}
where \key{parseability} is the amount of information that utterances provide about their underlying syntactic structures:
\begin{equation}
	R_{Pars} := I[\utterance,\tree] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)}
\end{equation}
and \key{predictability} is the negative entropy or surprisal of the language:
\begin{equation}
	R_{Pred} := - H[\utterance] = \sum_{u} p(u) \log p(u)
\end{equation}

We now discuss how this metric relates to other variants that have been proposed in the information-theoretic literature on language, and describe why~(\ref{eq:eff-general}-\ref{eq:efficiency-derived}) is most appropriate to our setting.
This efficiency metric~(\ref{eq:eff-general}-\ref{eq:efficiency-derived}) is equivalent to a deterministic version \cite{strouse2017deterministic} of the Information Bottleneck approach, which has been successfully applied to modeling word meaning across different domains \cite{zaslavsky2018efficient, zaslavsky2019semantic}.
The Information Bottleneck models complexity using a mutual information term instead of the entropy appearing in~(\ref{eq:eff-general}).
This mutual information formalization is appropriate to codes that are nondeterministic \cite{zaslavsky2018efficient}.
As we assume deterministic grammars that transduce every underlying syntactic structure into one surface order, our model corresponds to the Deterministic Information Bottleneck~\cite{strouse2017deterministic}, which arises when the encoder is deterministic, and which uses the entropy as the complexity measure.
Entropy also corresponds to the average of surprisal, which is a strong and linear predictor of human language processing effort~\cite{hale2001probabilistic,levy2008expectation,smith2013effect}.

A few other approaches share the mutual information term for informativity in (\ref{eq:eff-general}), while using complexity measures that are not explicitly information-theoretic.
%A few models have used other complexity measures that are not explicitly information-theoretic quantities.
In \cite{regier2007color, xu2014numeral, xu2016historical}, the complexity function is the number of different forms; in \cite{kemp2012kinship} it is the complexity of defining the concepts that are encoded.
The number of forms is not applicable in our case, as languages will typically have infinitely many sentences.
Notably, the models in \cite{regier2007color, xu2016historical} have since been reformulated successfully in the Information Bottleneck formalism \cite{zaslavsky2018efficient, zaslavsky2019semantic}, bringing them even closer to our formalization of efficiency.

In addition to these models, which quantify the efficiency of communication systems, there is closely related work formalizing the optimal choice of specific utterances in context, such as models of pragmatic reasoning in reference games \cite{frank2012predicting,goodman2013knowledge, kao2014nonliteral}.
In line with the other models, these assume that rational speakers choose utterances to optimize informativity about the referent object, and trade this off with the cost of the utterance, which is partly chosen to be the surprisal of the utterance \cite{bennett2018extremely}.

Beyond the study of natural language, the eficiency metric (\ref{eq:eff-general}) is also closely related to information-theoretic models in other disciplines.
The tradeoff between informativity and complexity of communication systems is studied by Rate-distortion theory \cite{shannon1948mathematical, berger1971rate}.
The \emph{Infomax Principle} derives parsimonious data representations by maximizing the mutual information between data and representations, subject to constraints on the representations~\cite{linsker1988self}; a constraint on the representation entropy leads to a metric equivalent to (\ref{eq:eff-general}) and to a version of the Free-Energy principle (see Section S3 in \citet{friston2010free}).

@@

%Other metrics very similar to~(\ref{eq:eff-general}) have been proposed in other computational work formalizing Zipf's ideas~\cite{frank2012predicting,zaslavsky2018efficient,kemp2012kinship,regier2015word,goodman2013knowledge,piantadosi2012communicative}.
%The idea of formalizing informativity using the mutual information $I[S, R]$ is common to all these approaches.

@@

The reason is that greater values of $\lambda$ can mathematically result in degenerate solutions.
To show this, note that the following inequality always holds:
\begin{equation}
I[\utterance; \tree] \leq H[\utterance].
\end{equation}
Therefore, if $\lambda \geq 1$, the efficiency objective satisfies 
%\begin{equation}
$R_{\textit{Eff}} = I[\utterance; \tree] - \lambda H[\utterance] \leq 0$,
%\end{equation}
and it takes the maximally possible value zero if there is only a single utterance $\utterance$, in which case both $I[\utterance; \tree]$ and $H[\utterance]$ are zero.
This is a degenerate language with only a single utterance, which is simultaneously used to convey all meanings.
While the design of our word order grammars (see \emph{Materials and Methods}) precludes a collapse of all syntactic structures to a single utterance, this shows that an objective with $\lambda \geq 1$ cannot be a generally applicable description of the efficiency of communication systems.
In conclusion, $\lambda$ is constrained to be in $[0,1)$, with values closer to $1$ placing similar weights on both predictability and parseability, whereas values closer to $0$ diminish the role of predictability.

@@

%- number of different forms 
%
%- complexity of the mapping 
%
%- MI between
%
%
%
%We model the first term as the 
% degree to which listeners can reconstruct dependency parse structures from an utterance, i.e., the \key{parseability} if the language. This is formalized as the amount of information that utterances $u$ provide about their underlying syntactic structures $t$:
%\begin{equation}
%	R_{Pars} := I[\utterance,\tree] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)}
%\end{equation}
%where the sum runs over all possible pairs of sentences $l$ and syntactic structures $t$ in the language.
%%This quantity describes the degree to which syntactic structures can be unambiguously recovered from sentences, and quantifies how successful communication is (cf. Figure~\ref{fig:comm}).
%
%
%
%We formalize the complexity of a language as its entropy~\cite{ferreri2003least,ferrericancho2007global,futrell2017memory}.
%In our setting, this corresponds to the average word-by-word surprisal, the degree to which sentences are unpredictable from the general statistics of the language.
%Surprisal has been found to be a highly accurate and general predictor of human online processing difficulty \cite{hale2001probabilistic,levy2008expectation,smith2013effect}.
%%and can be justified  in terms of predictive coding theories of information processing in the brain \cite{friston2009predictive} as well as the general algorithmic complexity of language generation and comprehension \cite{li2008introduction}.
%
%In expectation over all utterances $u$ in a language, the negative surprisal describes the \key{predictability}, or negative entropy, of the utterances:
%\begin{equation}
%	R_{Pred} := - H[\utterance] = \sum_{u} p(u) \log p(u)
%\end{equation}
%where the sum runs over all possible sentences $u$ that belong to the language.
%%In keeping with Zipf's Force of Unification, this quantity describes how homogeneous the language is, i.e., it is larger if the distribution over sentences is concentrated on a smaller number of frequent sentences. 
%
%
%
%The efficiency of a language is the weighted combination of these two scoring functions~\cite{ferreri2003least,frank2012predicting,zaslavsky2018efficient,kemp2012kinship,regier2015word,goodman2013knowledge}:
%\begin{equation}\label{eq:efficiency}
%	R_{\textit{Eff}} := R_{\textit{Parseability}} + \lambda R_\textit{Pred}
%\end{equation}
%with an interpolation weight $\lambda \in [0,1)$.
%In all results from Study 2 reported here, we set $\lambda := 0.9$ in Equation~\ref{eq:efficiency}.
%See SI appendix section 5 for mathematical discussion of $\lambda$, and robustness to other choices.
%
%

%\section{Relation to the Information Bottleneck}
%
%Our efficiency objective (Equation~\ref{eq:efficiency}), to be maximized for word order grammars $L$, is restated below.
%\begin{equation}
%    \label{eq:efficiency-again}
%    J_T(L) = I[L(T); T] - \lambda H[L(T)].
%\end{equation}
%This equation can be seen as a special case of the objective function proposed in \citet{tishby1999information} for general lossy compression, based on rate-distortion theory \citep{cover2006elements,harremoes2007information}. In that theory, the goal is to derive an encoding $\hat{X}=L(X)$ of some variable $X$ which preserves the information in $X$ relevant to some third variable $Y$, while minimizing irrelevant information. The three variables $Y$,$X$, and $\hat{X}$ are assumed to form a Markov chain $Y \rightarrow X \underrightarrow{L} \hat{X}$. Given those variables, the information bottleneck objective to be maximized is given in Eq.~\ref{eq:ib}.
%\begin{equation}
%    \label{eq:ib}
%    J_{X,Y}(L) = I[\hat{X}; Y] - \lambda I[\hat{X}; X].
%\end{equation}
%In this equation, the joint distribution of $X$ and $Y$ is seen as fixed, and the conditional distribution $L=\hat{X}|X$ is optimized.
%
%In our setting we consider a language $L$ expressing trees $T$ as utterances $U$: $T \underrightarrow{L} U$. To make our formulation more parallel to the information bottleneck, we assume the utterances $U$ are passed through a channel from producer to comprehender, and received as $\hat{U}=U$. We have the Markov chain $T \underrightarrow{L} U \rightarrow \hat{U}$, where we are assuming the conditional channel distribution $\hat{U}|U$ is fixed and the conditional distribution $L=U|T$ is to be optimized. Then our objective is parallel to Eq.~\ref{eq:ib}:
%\begin{equation}
%    \label{eq:efficiency-kinda-ib}
%    J_T(L) = I[U;T] - \lambda I[U;\hat{U}]. % TODO should this be I[\hat{U};T]?
%\end{equation}
%Assuming the channel $U \rightarrow \hat{U}$ is noiseless, we have $U=\hat{U}$, and thus $I[U;\hat{U}]=I[U;U] = H[U]$. Then Eq.~\ref{eq:efficiency-kinda-ib} reduces to a form of Eq.~\ref{eq:efficiency-again}:
%\begin{equation*}
%    J_T(L) = I[U;T] - \lambda H[U].
%\end{equation*}
%
%Our efficiency objective is thus closely related to the information bottleneck; the difference is that we are optimizing the mapping from underlying meanings to representations (strings) assuming that these strings will be transmitted in a fixed noiseless channel.
%
%

@@

\begin{table}
\centering
\small{
\begin{tabular}{l||ll|lll|lll}
Language & Pred. (t) & Parse. (t) & \multicolumn{3}{|c|}{Pred. (Binomial)} & \multicolumn{3}{|c|}{Parseab. (Binomial)} \\ 
&  $p$ & $p$ &  Est. &CI & $p$ & Est. & CI & $p$  \\ \hline \hline
\input{../results/plane/per-language-statistics.tex}
\end{tabular}
}
\caption{Per-language results in Study 1. For each language, we show the following: (1) $p$-values obtained from a one-sided $t$-test, for the null that the mean predictability/parseability of random grammars is at least as high as that of the real grammar. (2) Results from one-sided binomial tests, for the null that the the real grammar is better than at most $50 \%$ of random grammars. In addition to the $p$-value, we report point estimate and 95\% confidence interval for the fraction of random grammars that have values below real grammars.}\label{fig:pareto-per-lang-stats}
\end{table}

@@

We obtained the posterior density plots in Table 2 (Main Paper) and in Figure \ref{tab:all-predictions-1}) by applying the logistic transformation ($x \mapsto \frac{1}{1+\exp(-x)}$) to the posterior samples of $\alpha_j$ (\ref{eq:mixed-effects}).

@@

We provide an example in Figure~\ref{fig:grammar-sample}, illustrating grammar parameters for the relations in Figure 3 of the main paper.

@@

We validated the extracted grammars by comparing the dominant orders of six syntactic relations that are also annotated in the World Atlas of Linguistic Structures (WALS, \cite{haspelmath2005world}).
Among the eight Greenbergian correlations that we were able to test, five are annotated in WALS.
In Table~\ref{tab:grammars-wals}, we compare our grammars with WALS on these five relations, and the verb-object relation.
WALS has data for 74\% of the entries\footnote{Serbian and Croatian are listed as a single language Serbian-Croatian in WALS. In the table, we compare those with the grammar we extracted for Croatian, noting that it fully agrees with the Serbian grammar.}, and lists a dominant order for 91\% of these.
The grammars we extracted from the corpora agree with WALS in 96~\%  of these cases.

Note that the grammatical formalism simplifies some aspects of the word order regularities of natural languages.
For instance, it does not represent cases where ordering varies between main and embedded clauses, as it does not condition ordering decisions on the larger context.
%Our method extracts the most frequent ordering and applies it throughout all 
More complex and powerful ordering grammar models have been proposed \cite{futrell2015experiments, wang2016galactic}; however, they have similar limitations, and for our purposes, the model adopted here has the advantage of being simple and interpretable.
%There currently is no grammatical formalism that is at the same time capable of representing the full range of word order regularities that are found, 
%\cite{futrell2015experiments}

>>>>

\section{Formalizing Comunicative Efficiency}
\subsection{Derivation and Relation to Other Work}

@@

Here we discuss how our formalization of communicative efficiency relates to formalizations that have been proposed in the information-theoretic literature on language.
Across the literature, the core idea is to maximize the \key{amount of information} that linguistic forms provide about meanings, while constraining \key{complexity and diversity} of forms:
\begin{equation}\label{eq:eff-basic}
\text{Informativity} - \lambda \cdot \text{Complexity}
\end{equation}
with some differences in the precise formalization of these two quantities \cite{ferreri2003least,ferrericancho2007global,kemp2012kinship,frank2012predicting,goodman2013knowledge,kao2014nonliteral,xu2014numeral,kirby2015compression,regier2015word,xu2016historical,futrell2017memory,zaslavsky2018efficient,bennett2018extremely, peloquin2019interactions,zaslavsky2019semantic}.

\paragraph{Derivation of our Formalization}
The basis for our precise formalization is the function proposed in \cite{ferreri2003least,ferrericancho2007global,futrell2017memory,peloquin2019interactions} as a general efficiency metric for communicative systems.
%The key idea is to maximize the \key{amount of information} that linguistic forms provide about meanings, while constraining \key{complexity and diversity} of forms. %; there are some differences in the formalization of complexity.
If $S$ denotes signals (e.g., words, sentences) and $R$ denotes their referents (e.g., objects in a reference game), then this efficiency metric takes the form (notation slightly varies across these publications):
\begin{equation}\label{eq:eff-general}
\operatorname{I}[S, R] - \lambda \cdot \operatorname{H}[S]
\end{equation}
where $\operatorname{I}[S, R]$ describes the \key{informativity} of the signals $S$ about their referents $R$, and $\operatorname{H}[S]$ describes the \key{complexity} of the communication system, and $\lambda \geq 0$ trades off the two aspects of efficiency.
While prior studies \cite{ferreri2003least, kemp2012kinship,regier2015word,zaslavsky2018efficient} mostly considered settings where the signals $S$ are individual words without further structure, the signals are entire sentences $\utterance$ in our setting.
The underlying messages $R$ which the speaker aims to convey are the syntactic structures $\tree$.
By the principle of compositionality \cite{frege1892sinn}, the meaning of a sentence is a function of the meanings of the parts and how they are combined.
The syntactic structure specifies how the meanings of words are combined; therefore, recovering the syntactic structure is a prerequisite to understanding a sentence correctly.
%The idea that efficient languages will permit the listener to infer grammati
Hence, substituting utterances $\utterance$ for signals $S$, and syntactic structures $\tree$ for underlying messages $R$, into (\ref{eq:eff-general}), we  arrive at the following efficiency metric for word order:
\begin{equation}\label{eq:efficiency-derived}
	R_{\textit{Eff}} := R_{\textit{Pars}} + \lambda \cdot R_\textit{Pred}
\end{equation}
where \key{parseability} is the amount of information that utterances provide about their underlying syntactic structures:
\begin{equation}
	R_{Pars} := \operatorname{I}[\utterance,\tree] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)}
\end{equation}
and \key{predictability} is the negative entropy or surprisal of the language:
\begin{equation}
	R_{Pred} := - \operatorname{H}[\utterance] = \sum_{u} p(u) \log p(u)
\end{equation}
Parseability $\operatorname{I}[\utterance,\tree]$ is higher if utterances provide more information about their underlying syntactic structure.
Due to the identity $\operatorname{I}[\utterance, \tree] = \operatorname{H}[\tree] - \operatorname{H}[\tree|\utterance]$, parseability is maximized if every utterance can be parsed unambiguously -- that is, if the listener's uncertainty about syntactic structures given received utterances, $\operatorname{H}[\tree|\utterance]$, is zero.
Predictability $- \operatorname{H}[\utterance]$ is higher if the distribution over utterances is concentrated on a few utterances, and is maximized if there is just a single utterance.
It is also equal to the negative average of surprisal, which is a strong and linear predictor of human language processing effort~\cite{hale2001probabilistic,levy2008expectation,smith2013effect}.

\paragraph{Relation to Formalizations in Semantic Typology}
We now discuss how this metric relates to other variants that have been proposed in the information-theoretic literature on language, and describe why~(\ref{eq:eff-general}-\ref{eq:efficiency-derived}) is most appropriate to our setting.
This efficiency metric~(\ref{eq:eff-general}-\ref{eq:efficiency-derived}) is equivalent to a deterministic version \cite{strouse2017deterministic} of the Information Bottleneck approach, which has been successfully applied to modeling word meaning across different domains by \citet{zaslavsky2018efficient} and \citet{zaslavsky2019semantic}.
The Information Bottleneck models complexity using a mutual information term instead of the entropy appearing in~(\ref{eq:eff-general}).
This mutual information formalization is appropriate to codes that are nondeterministic \cite{zaslavsky2018efficient}.
We assume deterministic grammars that transduce every underlying syntactic structure into one surface order; our model therefore corresponds to the deterministic version of the Information Bottleneck, which arises when the encoder is deterministic, and which uses the entropy as the complexity measure~\cite{strouse2017deterministic}.

A few other approaches share the mutual information term for informativity in (\ref{eq:eff-general}), while using complexity measures that are not explicitly information-theoretic.
In studies of semantic typology by \citet{regier2007color, xu2014numeral, xu2016historical}, the complexity function is the number of different forms; in \citet{kemp2012kinship} it is the complexity of defining the concepts that are encoded.
%The number of forms is not applicable in our case, as languages will typically have infinitely many sentences.
Notably, the models of \citet{regier2007color} and \citet{xu2016historical} have since been reformulated successfully in the Information Bottleneck formalism \cite{zaslavsky2018efficient, zaslavsky2019semantic}, bringing them even closer to our formalization of efficiency.

\paragraph{Relation to Formalizations of Language Evolution}
There are also related models of the evolution of languages.
Most related to our work, \citet{kirby2015compression} describe a model of language evolution balancing the pressure for languages to be informative about the intended meanings, while minimizing complexity.
Formally, their model studies a Bayesian language learner who infers a language $h$ from data $d$ according to $P(h|d) \propto P(d|h) P(h)$, where $P(h)$ defines a prior distribution over languages, and $P(d|h)$ is the likelihood of observed data $d$ under the grammar $h$, assuming that speakers produce utterances pragmatically.
The prior $P(h)$ favors less complex languages; the likelihood $P(d|h)$ favors languages that communicate meanings unambiguously.
We now show that this model instantiates the basic objective~(\ref{eq:eff-basic}). 
If the dataset $d$ consists of observed pairs $(t,f)$ of meanings $t$ and forms $f$, and the language $h$ defines a set of possible pairs $(t,f)$, then the log-likelihood as defined by their model can be written as follows (up to constants)\footnote{We assume for simplicity that the error probability $\epsilon$ in the model is equal to $0$.}
\begin{equation}
\log P(d|h) = \sum_{(t,f) \in d} \log P(f|h,t) = \sum_{(t,f) \in d} \log P(f|h,t) \propto \sum_{(t,f) \in d} \log \frac{1}{|\{t' : (t', f) \in h\}|} = \sum_{(t,f) \in d} \log P(t|f)
\end{equation}
where $P(t|f)$ is the probability that the observed form $f$ referred to meaning $t$, as the model assumes uniform meaning distributions and uniform choice of appropriate forms.
Replacing the sum over the dataset $d$ by the expectation over the idealized full distribution over meaning-form pairs, this can be rewritten as 
\begin{equation}
- \operatorname{H}[t|f] = \operatorname{I}[t, f] - \operatorname{H}[t]
\end{equation}
where the first term is the mutual information between forms and meanings, as in our efficiency metric~(\ref{eq:eff-general}-\ref{eq:efficiency-derived}).
The second term, the entropy of meanings, is a constant independent of the form-meaning mapping.
The overall log probablity assigned by the Bayesian learner thus comes out to (up to constants)
\begin{equation}
\log P(h|d) = \operatorname{I}[t, f] + \lambda \log P(h)
\end{equation}
where the prior $P(h)$ favors simpler languages.
This result shows that the model of \citet{kirby2015compression} predicts that language evolution favors languages optimizing a function of the form~(\ref{eq:eff-basic}), with a informativity term identical to that of our model~(\ref{eq:eff-general}-\ref{eq:efficiency-derived}).

\paragraph{Relation to Formalizations of Pragmatics}
In addition to these models, which model the efficiency and evolution of communication systems, there is closely related work formalizing the optimal choice of specific utterances in context.
Our work is most closely related to the Rational Speech Acts model of pragmatic reasoning \cite{frank2012predicting,goodman2013knowledge, kao2014nonliteral}.
In line with the other models discussed here, it assumes that rational speakers choose utterances to optimize informativity about the referent object, and trade this off with the cost of the utterance, which is partly chosen to be the surprisal of the utterance \cite{bennett2018extremely, peloquin2019interactions}.
\citet{peloquin2019interactions} provide further discussion of the links between pragmatics and the efficiency metric (\ref{eq:eff-general}-\ref{eq:efficiency-derived}).

\paragraph{Relation to Models in Other Disciplines}
Beyond the study of natural language, the eficiency metric (\ref{eq:eff-general}) is also closely related to information-theoretic models in other disciplines.
The tradeoff between informativity and complexity of communication systems is studied by Rate-distortion theory \cite{shannon1948mathematical, berger1971rate}.
The \emph{Infomax Principle} derives parsimonious data representations by maximizing the mutual information between data and representations, subject to constraints on the representations~\cite{linsker1988self}; a constraint on the representation entropy leads to a metric equivalent to (\ref{eq:eff-general}) and to a version of the Free-Energy principle (see Section S3 in \citet{friston2010free}).

@@

%Other metrics very similar to~(\ref{eq:eff-general}) have been proposed in other computational work formalizing Zipf's ideas~\cite{frank2012predicting,zaslavsky2018efficient,kemp2012kinship,regier2015word,goodman2013knowledge,piantadosi2012communicative}.
%The idea of formalizing informativity using the mutual information $\operatorname{I}[S, R]$ is common to all these approaches.

@@

The reason is that greater values of $\lambda$ can mathematically result in degenerate solutions.
To show this, note that the following inequality always holds:
\begin{equation}
\operatorname{I}[\utterance; \tree] \leq \operatorname{H}[\utterance].
\end{equation}
Therefore, if $\lambda \geq 1$, the efficiency objective satisfies 
%\begin{equation}
$R_{\textit{Eff}} = \operatorname{I}[\utterance; \tree] - \lambda \operatorname{H}[\utterance] \leq 0$,
%\end{equation}
and it takes the maximally possible value zero if there is only a single utterance $\utterance$, in which case both $\operatorname{I}[\utterance; \tree]$ and $\operatorname{H}[\utterance]$ are zero.
This is a degenerate language with only a single utterance, which is simultaneously used to convey all meanings.
While the design of our word order grammars (see \emph{Materials and Methods}) precludes a collapse of all syntactic structures to a single utterance, this shows that an objective with $\lambda \geq 1$ cannot be a generally applicable description of the efficiency of communication systems.
In conclusion, $\lambda$ is constrained to be in $[0,1)$, with values closer to $1$ placing similar weights on both predictability and parseability, whereas values closer to $0$ diminish the role of predictability.

@@

%- number of different forms 
%
%- complexity of the mapping 
%
%- MI between
%
%
%
%We model the first term as the 
% degree to which listeners can reconstruct dependency parse structures from an utterance, i.e., the \key{parseability} if the language. This is formalized as the amount of information that utterances $u$ provide about their underlying syntactic structures $t$:
%\begin{equation}
%	R_{Pars} := \operatorname{I}[\utterance,\tree] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)}
%\end{equation}
%where the sum runs over all possible pairs of sentences $l$ and syntactic structures $t$ in the language.
%%This quantity describes the degree to which syntactic structures can be unambiguously recovered from sentences, and quantifies how successful communication is (cf. Figure~\ref{fig:comm}).
%
%
%
%We formalize the complexity of a language as its entropy~\cite{ferreri2003least,ferrericancho2007global,futrell2017memory}.
%In our setting, this corresponds to the average word-by-word surprisal, the degree to which sentences are unpredictable from the general statistics of the language.
%Surprisal has been found to be a highly accurate and general predictor of human online processing difficulty \cite{hale2001probabilistic,levy2008expectation,smith2013effect}.
%%and can be justified  in terms of predictive coding theories of information processing in the brain \cite{friston2009predictive} as well as the general algorithmic complexity of language generation and comprehension \cite{li2008introduction}.
%
%In expectation over all utterances $u$ in a language, the negative surprisal describes the \key{predictability}, or negative entropy, of the utterances:
%\begin{equation}
%	R_{Pred} := - \operatorname{H}[\utterance] = \sum_{u} p(u) \log p(u)
%\end{equation}
%where the sum runs over all possible sentences $u$ that belong to the language.
%%In keeping with Zipf's Force of Unification, this quantity describes how homogeneous the language is, i.e., it is larger if the distribution over sentences is concentrated on a smaller number of frequent sentences. 
%
%
%
%The efficiency of a language is the weighted combination of these two scoring functions~\cite{ferreri2003least,frank2012predicting,zaslavsky2018efficient,kemp2012kinship,regier2015word,goodman2013knowledge}:
%\begin{equation}\label{eq:efficiency}
%	R_{\textit{Eff}} := R_{\textit{Parseability}} + \lambda R_\textit{Pred}
%\end{equation}
%with an interpolation weight $\lambda \in [0,1)$.
%In all results from Study 2 reported here, we set $\lambda := 0.9$ in Equation~\ref{eq:efficiency}.
%See SI appendix section 5 for mathematical discussion of $\lambda$, and robustness to other choices.
%
%

%\section{Relation to the Information Bottleneck}
%
%Our efficiency objective (Equation~\ref{eq:efficiency}), to be maximized for word order grammars $L$, is restated below.
%\begin{equation}
%    \label{eq:efficiency-again}
%    J_T(L) = \operatorname{I}[L(T); T] - \lambda \operatorname{H}[L(T)].
%\end{equation}
%This equation can be seen as a special case of the objective function proposed in \citet{tishby1999information} for general lossy compression, based on rate-distortion theory \citep{cover2006elements,harremoes2007information}. In that theory, the goal is to derive an encoding $\hat{X}=L(X)$ of some variable $X$ which preserves the information in $X$ relevant to some third variable $Y$, while minimizing irrelevant information. The three variables $Y$,$X$, and $\hat{X}$ are assumed to form a Markov chain $Y \rightarrow X \underrightarrow{L} \hat{X}$. Given those variables, the information bottleneck objective to be maximized is given in Eq.~\ref{eq:ib}.
%\begin{equation}
%    \label{eq:ib}
%    J_{X,Y}(L) = \operatorname{I}[\hat{X}; Y] - \lambda \operatorname{I}[\hat{X}; X].
%\end{equation}
%In this equation, the joint distribution of $X$ and $Y$ is seen as fixed, and the conditional distribution $L=\hat{X}|X$ is optimized.
%
%In our setting we consider a language $L$ expressing trees $T$ as utterances $U$: $T \underrightarrow{L} U$. To make our formulation more parallel to the information bottleneck, we assume the utterances $U$ are passed through a channel from producer to comprehender, and received as $\hat{U}=U$. We have the Markov chain $T \underrightarrow{L} U \rightarrow \hat{U}$, where we are assuming the conditional channel distribution $\hat{U}|U$ is fixed and the conditional distribution $L=U|T$ is to be optimized. Then our objective is parallel to Eq.~\ref{eq:ib}:
%\begin{equation}
%    \label{eq:efficiency-kinda-ib}
%    J_T(L) = \operatorname{I}[U;T] - \lambda \operatorname{I}[U;\hat{U}]. % TODO should this be \operatorname{I}[\hat{U};T]?
%\end{equation}
%Assuming the channel $U \rightarrow \hat{U}$ is noiseless, we have $U=\hat{U}$, and thus $\operatorname{I}[U;\hat{U}]=\operatorname{I}[U;U] = \operatorname{H}[U]$. Then Eq.~\ref{eq:efficiency-kinda-ib} reduces to a form of Eq.~\ref{eq:efficiency-again}:
%\begin{equation*}
%    J_T(L) = \operatorname{I}[U;T] - \lambda \operatorname{H}[U].
%\end{equation*}
%
%Our efficiency objective is thus closely related to the information bottleneck; the difference is that we are optimizing the mapping from underlying meanings to representations (strings) assuming that these strings will be transmitted in a fixed noiseless channel.
%
%

@@

\begin{table}
\centering
\small{
\begin{tabular}{l||ll|lll|lll}
Language & Pred. (t) & Parse. (t) & \multicolumn{3}{c|}{Pred. (Binomial)} & \multicolumn{3}{c}{Parseab. (Binomial)} \\ 
&  $p$ & $p$ &  Est. &CI & $p$ & Est. & CI & $p$  \\ \hline \hline
\input{../results/plane/per-language-statistics.tex}
\end{tabular}
}
\caption{Per-language results in Study 1. For each language, we show the following: (1) $p$-values obtained from a one-sided $t$-test, for the null that the mean predictability/parseability of random grammars is at least as high as that of the real grammar. (2) Results from one-sided binomial tests, for the null that the the real grammar is better than at most $50 \%$ of random grammars. In addition to the $p$-value, we report point estimate and 95\% confidence interval for the fraction of random grammars that have values below real grammars.}\label{fig:pareto-per-lang-stats}
\end{table}

@@

We obtained the posterior density plots in Table 2 (Main Paper) and in Figure (\ref{tab:all-predictions-1}) by applying the logistic transformation ($x \mapsto \frac{1}{1+\exp(-x)}$) to the posterior samples of $\alpha_j$ (\ref{eq:mixed-effects}).

@@

We provide an example in Figure~\ref{fig:grammar-sample}, illustrating grammar parameters for the relations in Figure 3 of the main paper.
Note that the grammatical formalism simplifies some aspects of the word order regularities of natural languages.
For instance, it does not represent cases where ordering varies between main and embedded clauses, as it does not condition ordering decisions on the larger context.
%Our method extracts the most frequent ordering and applies it throughout all 
More complex and powerful ordering grammar models have been proposed \cite{futrell2015experiments, wang2016galactic}; however, they have similar limitations, and for our purposes, the model adopted here has the advantage of being simple and interpretable.
%There currently is no grammatical formalism that is at the same time capable of representing the full range of word order regularities that are found, 
%\cite{futrell2015experiments}

@@

We validated the extracted grammars by comparing the dominant orders of six syntactic relations that are also annotated in the World Atlas of Linguistic Structures (WALS, \cite{haspelmath2005world}).
Among the eight Greenbergian correlations that we were able to test, five are annotated in WALS: adpositions, complementizers, relative clauses, genitives, and oblique PPs.
In Table~\ref{tab:grammars-wals}, we compare our grammars with WALS on these five relations, and the verb-object relation.
WALS has data for 74\% of the entries\footnote{Serbian and Croatian are listed as a single language Serbian-Croatian in WALS. In the table, we compare those with the grammar we extracted for Croatian, noting that it fully agrees with the Serbian grammar.}, and lists a dominant order for 91\% of these.
The grammars we extracted from the corpora agree with WALS in 96~\%  of these cases.

\begin{table}
\small{
\begin{center}
\begin{tabular}{l||ll|ll|ll|ll|ll|ll|llllll}
%	   for(dep in c("obj", "lifted_mark", "acl", "nmod", "obl")) {
		   Language 
		   &	\multicolumn{2}{c|}{Objects} 
		   &	\multicolumn{2}{c|}{Adpositions} 
		   &	\multicolumn{2}{c|}{Compl.} 
		   &	\multicolumn{2}{c|}{Rel.Cl.} 
		   &	\multicolumn{2}{c|}{Genitive} 
		   &	\multicolumn{2}{c|}{PP}  \\ \hline\hline
	\input{../results/grammars/comparison-table.tex}
\end{tabular}
\end{center}
}
\caption{Comparing grammars extracted from databases to linguistic judgments in the World Atlas of Linguistic Structures. For each of the six syntactic relation, the first column provides the ordered coded in the extracted grammar; the second column provides the order coded in WALS (DH for dependent-head, HD for head-dependent order). `?' indicates that WALS has no data.
$*$ indicates that WALS does not list a dominant order; as \citet{dryer2011evidence} describes, this can mean that neither order is dominant in the language, or that insufficient data was available when compiling WALS.
Finally, `--' indicates that the relation does not occur in the Universal Dependencies corpus.
}\label{tab:grammars-wals}
\end{table}
